---
title: "Automatic Sampling and Analysis of YouTube Data"
subtitle: "Tools for collecting YouTube data"
author: "Johannes Breuer, Annika Deubel, & M. Rohangis Mohseni"
date: "February 14th, 2023"
presenter: Everybody

---

layout: true

```{r child = "./content/config/sessions_setup.Rmd"}
```

---

## How to collect *YouTube* data

There are many different ways in which data from *YouTube* and other social media can be collected (see [Breuer et al., 2020](https://journals.sagepub.com/doi/10.1177/1461444820924622)):

- Manually (e.g., via copy & paste and manual content analysis)

- Using existing data, such as [*YouNiverse: Large-Scale Channel and Video Metadata from English YouTube*](https://zenodo.org/record/4650046) (also see the accompanying preprint by [Ribeiro & West, 2021](https://arxiv.org/abs/2012.10378))

- Automatically via the *YouTube* API or web scraping

---

## Comparisons of Approaches for Collecting *YouTube* Data

.small[
```{r tools table 1, echo = F}
tools <- data.frame(
  "Software" = c("YouTube Data Tools 1.22", "Webometric 4.1", "Tuber 0.9.9", "vosonSML 0.29.13", "youtubecaption 1.0.0"),
  "Type" = c("Website", "Standalone app", "R package", "R package", "R package"),
  "Can collect" = c("Channel Info, Video Info, Comments", "Channel Info, Video Info, Comments, Video Search", "Channel Info, Video Info, Comments, Subtitles, All searches", "Video IDs, Comments", "Subtitles"),
  "Comment Scope" = c("Only all", "100 most recent or all", "20-100 most recent or all", "1-x top-level", "n/a"),
  "Needs API Key" = c("No", "Yes", "Yes", "Yes", "No")
  , stringsAsFactors = FALSE
  , check.names = FALSE
)
knitr::kable(
  tools
  , format = "html"
  , align = "cc"
  , escape = FALSE
)
```
]

---

## Exemplary Comparison of the Different Tools

.small[
```{r tools table 2, echo = F}
usability <- data.frame(
"Software" = c("YouTube Data Tools 1.22", "Webometric 4.1", "Tuber 0.9.9", "vosonSML 0.29.13"),
  "Ease of Use" = c("High", "Low", "Low", "Low"),
  "Disadvantages" = c("Lacking flexibility, less information", "Only first 5 follow-up comments, no error feedback, undetectable time-outs", "Only first 5 follow-up comments", "Lacking flexibility, only comments"),
  "No. of Comments" = c("52,243", "49,150", "49,139", "50,619")
  , stringsAsFactors = FALSE
  , check.names = FALSE
)
knitr::kable(
  usability
  , format = "html"
  , align = "cc"
  , escape = FALSE
)
```
]

Example data source: [Dayum Video](https://www.youtube.com/watch?v=DcJFdCmN98s)

---

## A note on using FOSS

The tools listed before are free and open source software (FOSS). Using FOSS has many advantages (availability, adaptability, etc.). However, one risk associated with using FOSS is that tools are not maintained anymore and, hence, cease to function. After all, people create and maintain these tools in their spare time or as side projects and this work is often not recognized enough (esp. within academia). For this reason it is especially important to acknowledge the work that goes into these tools by properly citing them.

.small[
```{r cite-pkg, echo=TRUE}
citation("tuber")
```
]

---

class: center, middle

# Any questions so far?